Deep recurrent neural networks for supernovae classification

### 2017-02-06 paper/sn_submission_v3/Referee reply.txt

We wish to thank the referee for their points and we have done our best to expand on each of them fully.

@@

(1) The work as presented is in almost acceptable form. The part about using just the
first few points is powerful and will be welcomed by many. If using the entire
light-curve during testing provides somewhat worse results, then perhaps there
is a way to improve the model even further.

There could be possibly ways to improve the model further, but this is not a certainty. One reason that improving the model further may not actually help is that, due to the way RNNs work, including the information from after the early epoch challenge skews the overall weights of the network meaning that the early epoch weights are less well known compared to training using only the early epoch challenge data. Since this is mentioned in the text we have not expanded further.

@@

(6a) In Table 2 please include the 0.052% even if the results are bad.

They are already in the table in every case - 1,045 is equivalent to a train fraction of 0.052 (not 0.052%).

(6b) Are the early-epoch results for SN1a or 123? Can the results be provided for both sets?

The early-epoch challenge is for SN1a only and not 123. We have given 123 a go and added the results to the table. The difference between 123 categorisation using the full light-curve and the early-epoch data is very similar to the difference between the SN1a full light-curve and early-epoch data. The results are not as good, but still surprisingly high considering the length of the data.

@@

(6c) In the middle section, in the last 2 lines, Completeness seems to rise when Host
information is dropped. Thats anomalous. If you understand why, a comment will
be useful. Otherwise those are the kind of error-bars that become meaningful:
are both 73+/-2%? You can still fit the table on a single page by abbreviating
contents in the first column.

This was a statistical fluctuation. We have added the errors on the purity and completeness. As it can be seen the errors on both of the bottom lines of the second section are ~4%, approximately the same as the difference between the means. The same is true for the 123 categorisation in the early-epoch challenge where the completeness without host-z is higher than with host-z, but the error on the completeness is also very large.

(7) It is understandable that your methods are not comparable with various methods
used during SPCC, but perhaps its still worth mentioning the best numbers
there?

We have added a paragraph stating the numbers of the highest average F_1 method (Sako) from the SPCC results paper. We have also said that they other methods performed similarly but they cannot be compared well to our results.

@@

(9) The following is a loose sentence: "Moreover, we have ...trained on the whole
light-curve". Especially so since you do not normalize time to the peak value so the
starting point could be arbitrary (or is your padding mitigating exactly that?
if so, clarify a little). Otherwise quantify (at least verbally) what you mean
by the above sentence.

We mean that training using the data from the early epoch challenge gives a better early-time predictor than training using the whole light-curve. We have rewritten the sentence to make it clearer.

(10) You mention that perhaps even raw images could be used etc., but you have not
even discussed the effects of different filters etc. What happens if you use
just one or two filters, for instance? So, either go into more details, or
exclude pure speculation.

This isn't pure speculation, this is a consideration for further work. It is possible to data mine images and pipe this into an RNN. We could do this for a single filter or we could do this for many filters. If one filter was missing then the input could be augmented as done here and the other raw images convolved to extract features. The principal is identical to what is done here, but with a huge amount more information. We have not changed the text here as it is an interesting prospect which others (as well as ourselves) might like to consider, but would need considerable work to go into detail.

### 2017-02-06 paper/sn_submission_v3/ms.tex
v3 Paper submission

 Each neuron is connected to another by a weight matrix, and the optimal weights are found by back-propagating the errors from a {\em loss function} of the output layer. For classification problems, this is typically the categorical cross-entropy between predictions and targets, defined as 
\begin{equation}
L= -\sum_{i,j} t_{i,j} \log \left( p_{i,j} \right)
\end{equation}
where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities. Back-propogation takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.

@@

In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, consisting of 21319 simulated supernove light curves.  Each supernovae sample consists of a time series of flux measurements, with errors, in the $g,r,i,z$ bands (one band for each timestep), along with the position on the sky and dust extinction. An example set of light curves is shown in Fig.~\ref{fig:lightcurve}.

@@

In Fig.~\ref{fig:violinplot} we show how the length of the grouped-time data vector is related to the duration of the light-curve. The bottom left subplot shows that more total number of day since the beginning of observation of the light-curve results in a greater number of grouped time elements in the vector. The upper subplot shows the distribution of observation lengths in the SPCC data varies significantly with two distinct peaks. These are grouped into an average of 40-3element data vectors as can be seen in the bottom right subplot.

@@

It is difficult to directly compare the results from the SPCC directly with this work, but using post-SPCC data, we obtain results which are competitive with previous approaches. The analyses by~\cite{Karpenka:2012pm} and~\cite{Newling:2010bp} are easier to compare. Along with~\cite{Lochner:2016hbn} these employ a two-step process, where features are first extracted by various methods before machine learning classification. The results obtained for similar sized training sets are comparable as can be seen in the top section of Table~\ref{tab:05_SN1a}. When using half the dataset to train on we get a higher $F_1$ value, $F_1=0.64$ compared to $F_1=0.58$ in~\cite{Karpenka:2012pm}. The value in~\cite{Newling:2010bp} is also similar given that the sample size is smaller. For a smaller sample training set of 5.2\% of all the data we again perform similarly to~\cite{Karpenka:2012pm} but under perform compare to~\cite{Newling:2010bp} taking into account the slightly larger sample size in the latter case.

@@

>>>>

 Each neuron is connected to another by a weight matrix, and the optimal weights are found by back-propagating the errors from a {\em loss function} of the output layer. For classification problems, this is typically the categorical cross-entropy between predictions and targets, defined as 
\begin{equation}
L= -\sum_{i,j} t_{i,j} \log \left( p_{i,j} \right)
\end{equation}
where $i,j$ run over the class labels, $t_{i,j}$ are the targets for each class (either 0 or 1) and $p_{i,j}$ are the predicted probabilities. Back-propagation takes the derivative of the loss with respect to the weights $W$ of the output layer, $\partial L/\partial W$, and uses the chain rule to update the weights in the network.

@@

In this paper we will consider data from the Supernovae Photometric Classification Challenge (SPCC)~\cite{Kessler:2010wk,Kessler:2010qj}, consisting of 21319 simulated supernova light curves.  Each supernovae sample consists of a time series of flux measurements, with errors, in the $g,r,i,z$ bands (one band for each timestep), along with the position on the sky and dust extinction. An example set of light curves is shown in Fig.~\ref{fig:lightcurve}.

@@

In Fig.~\ref{fig:violinplot} we show how the length of the grouped-time data vector is related to the duration of the light-curve. The bottom left subplot shows that more total number of day since the beginning of observation of the light-curve results in a greater number of grouped time elements in the vector. The upper subplot shows the distribution of observation lengths in the SPCC data varies significantly with two distinct peaks. These are grouped into an average of 40-element data vectors as can be seen in the bottom right subplot.

@@

It is difficult to directly compare the results from the SPCC challenge in~\cite{Kessler:2010qj} with this work since the figure of merit is quoted as a function of redshift and a non-representative set of light-curves was originally used. In~\cite{Kessler:2010qj} the method of~\cite{2008AJ....135..348S} had the highest average $F_1\sim0.79$, with 76\% purity and the 96\% completeness. This is a, somewhat, confusing average as $F_1\sim 0.4$ at a redshift $z\sim0.1$ up to $F_1\sim1$ at $z\sim0.9$. Other methods performed similarly.

It is better to consider comparison with other methods using post-SPCC data, for we obtain results which are competitive with previous approaches. The analyses by~\cite{Karpenka:2012pm} and~\cite{Newling:2010bp} are easier to compare. Along with~\cite{Lochner:2016hbn} these employ a two-step process, where features are first extracted by various methods before machine learning classification. The results obtained for similar sized training sets are comparable as can be seen in the top section of Table~\ref{tab:05_SN1a}. When using half the dataset to train on we get a higher $F_1$ value, $F_1=0.64$ compared to $F_1=0.58$ in~\cite{Karpenka:2012pm}. The value in~\cite{Newling:2010bp} is also similar given that the sample size is smaller. For a smaller sample training set of 5.2\% of all the data we again perform similarly to~\cite{Karpenka:2012pm} but under perform compare to~\cite{Newling:2010bp} taking into account the slightly larger sample size in the latter case.

@@

\bibitem[Sako et al.(2008)]{2008AJ....135..348S} Sako, M., Bassett, B., Becker, A., et al.\ 2008, \aj, 135, 348-373
